---
title: "Inequality in the United States"
author: "Ari Boyarsky"
date: "October 4, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Using R and GGPlot to Map Inequality

```{r, echo=TRUE, message=FALSE}
rm(list = ls())


US_GINI_Data <- read.csv("US Census Data Intra-US GINI/ACS_15_1YR_B19083_with_ann.csv")

suppressMessages(library(rgeos))
library(ggplot2)
library(rgeos)
library(plyr)
library(scales)
require(graphics)
library(ggthemes)
library(colorspace)

# read csv
US_GINI_Data <- read.csv("US Census Data Intra-US GINI/ACS_15_1YR_B19083_with_ann.csv")

# state data
states_map <- map_data("state")

# Let's give better names
colnames(US_GINI_Data) <- c("GEO.id", "GEO.id2", "state", "GINI", "GINI2")

# change to lower case
gini <- as.data.frame(lapply(US_GINI_Data, tolower), stringsAsFactors = FALSE)
# we have to trasform GINI to num
gini <- transform(gini, GINI = as.numeric(GINI))

# megre map data and GINI data
gini_map <- merge(states_map, gini, by.x = "region", by.y = "state")

# create map, g
g <- ggplot(data=gini_map, mapping = aes(x=long, y=lat, group=group))+
  geom_polygon(aes(fill=GINI))+
  geom_path() + coord_map("mercator")

remove_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
)
# add title
g <- g+remove_axes+ggtitle('GINI Accross the United States')
# change data color encoding
g <- g+scale_fill_distiller(name="GINI", palette = "YlGn", breaks = pretty_breaks(n = 5))

# print graph
g
```

The above choropleth shows the GINI level within the United States, differentiated by State. The lighter the color fill, the higher the GINI. GINI is measured on a 0 to 1 scale, where 0 is a theoretically equal income distribution. While 1 represents a society where all the income is controlled by a single holder. 

We can use R to gain more information on the GINI distribution between U.S. States, for example:

```{r, echo=TRUE}
summary(gini)
```
This analysis shows us the minimum U.S. State GINI, as well as the maximum. Furthermore, we can see the mean U.S. GINI is `0.4675`, while the median is `0.466`. Of course, this is different from the U.S. GINI coefficent. 

## References

U.S. Census Bureau, 2014 American Community Survey 1-Year Estimates